---
title: "Google Routing"
output:
  html_document:
    keep_md: yes
date: "2023-12-20"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(googleway)
library(tidyverse)
library(ggmap)
library(sf)
library(sfext)
library(googlePolylines)
library(mapview)
googleway::set_key(key = map_key)
ggmap::register_google(map_key)
```

https://stackoverflow.com/questions/59775761/drawing-driving-routes-in-r

# Setup data frame to be filled in

Cities and stations

1. Ottawa
    * Fallowfield Station
    * Millennium / Trim
2. Montreal 
    * Brossard Station
    * Rivière-des-Prairies Station
3. Toronto
    * Vaughan Metropolitan Centre Stations
    * Scarborough GO Station
4. Vancouver
    * Surrey Central Station
    * Edgemont Village Exchange
5. Edmonton
    * Heritage Valley Station
    * Nakî Transit Centre & Park and Ride
6. Calgary
    * North Pointe Transit Terminal
    * McKenzie Towne

```{r}
city <- c("Ottawa", "Montreal", "Toronto", "Vancouver", "Edmonton", "Calgary")
stations <- c("Fallowfield Station, Ottawa, ON", "Millennium / Trim Station, Ottawa, ON", "Brossard Station, Montreal, QC", "Rivière-des-Prairies Station, Montreal, QC", "Vaughan Metropolitan Centre Stations, Toronto, ON", "Scarborough GO Station, Toronto, ON", "Surrey Central Station, Vancouver, BC", "Edgemont Village Exchange, Vancouver, BC", "Heritage Valley Station, Edmonton, AB", "Nakî Transit Centre & Park and Ride, Edmonton, AB", "North Pointe Transit Terminal, Calgary, AB", "McKenzie Towne, Calgary, AB")
```

# Ottawa
## Fallowfield Station

### Draw buffer and select points

```{r}
departure <- "Fallowfield, Ottawa, Canada"

map <- get_googlemap(departure)

stations <- geocode(location = stations, output = "more", source = "google")

stations_sf <- stations %>%
                st_as_sf(coords = c("lon", "lat"), crs = 4326)

stations_buffer = st_buffer(stations_sf, 5000)
```

#### Interactive map view of buffers

```{r}
mapview(stations_buffer)
```

#### Generate points within buffer

```{r}
points_sample_sf <- st_sample(stations_buffer, size = c(25,25), type = "random")
points_sample <- sf_to_df(points_sample_sf)
```

#### Check to see if they worked

```{r}
ottawa_check_buffer <- stations_buffer %>% slice(1:2)

plot(st_geometry(ottawa_check_buffer))
plot(points_sample_sf, pch = 20, add= TRUE)
```

Setup departure location and arrival time

```{r}
arrival_time <- as.POSIXct("2024-06-01 08:30:00", tz = "America/Toronto")
```

## Bicycle routes for Ottawa

```{r}
ottawa_bike <- google_directions(origin = "Fallowfield, Ottawa, Canada",
         destination = "Downtown, Ottawa, Canada",
         arrival_time =  arrival_time,
         mode = "bicycling",
         alternatives = FALSE,
         units = "metric")

ottawa_bike[[1]]
```

### Create a data frame and map for route

```{r}
ottawa_bike_df <- ottawa_bike$routes

ottawa_bike_lines <- decode_pl(ottawa_bike_df$overview_polyline$points)

ggmap(map) +
  geom_point(data = ottawa_bike_lines, aes(x = lon, y = lat)) 
```

Add in distance and time to the data final data frame

```{r}
distance_meters <- ottawa_bike[["routes"]][["legs"]][[1]][["distance"]][["value"]]
duration_seconds <- ottawa_bike[["routes"]][["legs"]][[1]][["duration"]][["value"]]
```

# Transit routes for Ottawa

```{r}
ottawa_transit <- google_directions(origin = "Fallowfield, Ottawa, Canada",
         destination = "Downtown, Ottawa, Canada",
         arrival_time =  arrival_time,
         mode = "transit",
         alternatives = FALSE,
         units = "metric")

ottawa_transit[[1]]
```

### Create a data frame and map for route

```{r}
ottawa_transit_df <- ottawa_transit$routes

ottawa_transit_lines <- decode_pl(ottawa_transit_df$overview_polyline$points)

map <- get_googlemap(departure)

ggmap(map) +
  geom_point(data = ottawa_transit_lines, aes(x = lon, y = lat)) 
```




## NOT RUN

```{}
df_locations<-structure(list(origin = c("WARWICK", "EAST PROVIDENCE", "WARREN", 
                                        "CENTERDALE", "CENTRAL FALLS", "DAVISVILLE", "NORTH PROVIDENCE", 
                                        "EAST PROVIDENCE", "PROVIDENCE", "CHEPACHET"), destination = c("CENTERDALE", "EAST PROVIDENCE", "BRISTOL", "JOHNSTON", "CRANSTON", "WARWICK","NORTH PROVIDENCE", "EAST PROVIDENCE", "WARREN", "CHEPACHET")), class = "data.frame", row.names = c(NA, -10L))

## loop over each pair of locations, and extract the polyline from the result
lst_directions <- apply(df_locations, 1, function(x){
  res <- google_directions(
    origin = x[['origin']]
    , destination = x[['destination']]
  )
})

lst_directions[[2]][ , "geocoded_waypoints"] 

lst_directions <- apply(df_locations, 1, function(x){
  df_result <- data.frame(
    origin = x[['origin']]
    , destination = x[['destination']]
    , route = res$routes$overview_polyline$points
  )
  return(df_result)
})

df_directions <- do.call(rbind, lst_directions)

## plot the map
google_map() %>%
  add_polylines(data = df_directions, polyline = "route")
```

         
```{r, include = FALSE}
clear_keys()
```

