---
title: "Google Routing"
output:
  html_document:
    keep_md: yes
date: "2023-12-20"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(googleway)
library(tidyverse)
library(ggmap)
library(sf)
library(sfext)
library(googlePolylines)
library(mapview)
library(jsonlite)
library(collapse)
googleway::set_key(key = map_key)
ggmap::register_google(map_key)
```

# Ottawa

## Barrhaven Center Station

### Draw buffer and select points

```{r}
destination <- "National Gallery of Canada, Ottawa, Canada"        #### EDIT ME
```

### Set origin point

```{r}
origin <- "Barrhaven Center Station, Ottawa, Canada"    #### EDIT ME
```

#### Setup departure location and arrival time

June 22, 2024

```{r}
arrival_time <- as.POSIXct("2024-06-25 08:30:00", tz = "America/Toronto")       #### EDIT ME
```

```{r}
stations <- geocode(location = origin, output = "more", source = "google")

stations_sf <- stations %>%
                st_as_sf(coords = c("lon", "lat"), crs = 4326)

stations_buffer = st_buffer(stations_sf, 5000)

stations_buffer <- as_sf(stations_buffer)
```

#### Interactive map view of buffers

```{r}
mapview(stations_buffer)
```

#### Generate points within buffer

```{r}
set.seed(100)
points_sample_sf <- st_sample(stations_buffer, size = c(25,25), type = "random")
points_sample_sf <- st_as_sf(points_sample_sf)
points_sample <- sf_to_df(points_sample_sf)
```

#### Check to see if they worked

```{r}
ottawa_check_buffer <- stations_buffer %>% slice(1:2)

plot(st_geometry(ottawa_check_buffer))
plot(points_sample_sf, pch = 20, add= TRUE)
```

```{r}
points_sample$destination <- destination
points_sample$origin <- origin
points_sample$lat_lon <- paste(points_sample$lat, points_sample$lon)
```

#### Checking points and general mapping

```{r}
lst_directions <- apply(points_sample, 1, function(x){
  res <- google_directions(
    origin = x[['lat_lon']], 
    destination = x[['destination']]    
  )
df_result <- data.frame(
    origin = x[['lat_lon']], 
    destination = x[['destination']], 
    route = res$routes$overview_polyline$points 
)
return(df_result)
})

df_directions <- do.call(rbind, lst_directions)

google_map() %>%
  add_polylines(data = df_directions, polyline = "route") 
```


# BIKE: Getting cycling routes and points

```{r}
bike_directions <- apply(points_sample, 1, function(x){
  res <- google_directions(
    origin = x[['lat_lon']], 
    destination = x[['destination']],
    arrival_time =  arrival_time,
    mode = "bicycling",
    alternatives = FALSE,
    units = "metric"
  )
})
```

### Checking geocode and route status

```{r}
get_elem(bike_directions, "status", recursive = TRUE, DF.as.list = TRUE)
```

### Getting the distance data from bike

```{r}
distance_bike <- get_elem(bike_directions, "distance", recursive = TRUE, DF.as.list = TRUE)
distance_bike <- as_tibble(distance_bike)
distance_bike <- distance_bike %>% 
                            rename_with(~paste0("var", seq_along(.)))
distance_bike <- distance_bike %>%
                            filter(!row_number() %in% c(2))

distance_bike <- distance_bike %>% unnest(c(var1, var2, var3, var4, var5, var6, var7, var8, var9, var10, var11, var12, var13, var14, var15, var16, var17, var18, var19, var20, var21, var22), names_sep = "unique", names_repair = "universal")

distance_bike <- distance_bike %>%
                    select(contains("value"))

distance_bike <- distance_bike %>% 
                    pivot_longer(cols = starts_with("var"),
                                names_to = "route",
                                names_prefix = "wk",
                                values_to = "distance_meters",
                                values_drop_na = FALSE
                              )

distance_bike$distance_km <- distance_bike$distance_meters / 1000

ggplot(distance_bike) + 
  geom_histogram(aes(distance_km))

summary(distance_bike$distance_km)
```

### Getting duration data for bike

```{r}
duration_bike <- get_elem(bike_directions, "duration", recursive = TRUE, DF.as.list = TRUE)
duration_bike <- as_tibble(duration_bike)
duration_bike <- duration_bike %>% 
                            rename_with(~paste0("var", seq_along(.)))
duration_bike <- duration_bike %>%
                            filter(!row_number() %in% c(2))

duration_bike <- duration_bike %>% unnest(c(var1, var2, var3, var4, var5, var6, var7, var8, var9, var10, var11, var12, var13, var14, var15, var16, var17, var18, var19, var20, var21, var22), names_sep = "unique", names_repair = "universal")

duration_bike <- duration_bike %>%
                    select(contains("value"))

duration_bike <- duration_bike %>% 
                    pivot_longer(cols = starts_with("var"),
                                names_to = "route",
                                names_prefix = "wk",
                                values_to = "duration_seconds",
                                values_drop_na = FALSE
                              )

duration_bike$duration_hours <- duration_bike$duration_seconds / 3600

ggplot(duration_bike) + 
  geom_histogram(aes(duration_hours))

summary(duration_bike$duration_hours)
```

### Combining distance and duration for bike

```{r}
distance_duration_bike <- full_join(distance_bike, duration_bike)
```


# TRANSIT: Getting transit routes and points

```{r}
transit_directions <- apply(points_sample, 1, function(x){
  res <- google_directions(
    origin = x[['lat_lon']], 
    destination = x[['destination']],
    arrival_time =  arrival_time,
    mode = "transit",
    alternatives = FALSE,
    units = "metric"
  )
})

get_elem(transit_directions, "status", recursive = TRUE, DF.as.list = TRUE)
```

### Getting the distance data for transit

```{r}
distance_transit <- get_elem(transit_directions, "distance", recursive = TRUE, DF.as.list = TRUE)
distance_transit <- as_tibble(distance_transit)
distance_transit <- distance_transit %>% 
                            rename_with(~paste0("var", seq_along(.)))
distance_transit <- distance_transit %>%
                            filter(!row_number() %in% c(2))

distance_transit <- distance_transit %>% unnest(c(var1, var2, var3, var4, var5, var6, var7, var8, var9, var10, var11, var12, var13, var14, var15, var16, var17, var18, var19, var20, var21, var22), names_sep = "unique", names_repair = "universal")

distance_transit <- distance_transit %>%
                      select(contains("value"))

distance_transit <- distance_transit %>% 
                    pivot_longer(cols = starts_with("var"),
                                names_to = "route",
                                names_prefix = "wk",
                                values_to = "distance_meters",
                                values_drop_na = FALSE
                              )

distance_transit$distance_km <- distance_transit$distance_meters / 1000

ggplot(distance_transit) + 
  geom_histogram(aes(distance_km))

summary(distance_transit$distance_km)
```

### Getting duration data for transit

```{r}
duration_transit <- get_elem(transit_directions, "duration", recursive = TRUE, DF.as.list = TRUE)
duration_transit <- as_tibble(duration_transit)
duration_transit <- duration_transit %>% 
                            rename_with(~paste0("var", seq_along(.)))
duration_transit <- duration_transit %>%
                            filter(!row_number() %in% c(2))

duration_transit <- duration_transit %>% unnest(c(var1, var2, var3, var4, var5, var6, var7, var8, var9, var10, var11, var12, var13, var14, var15, var16, var17, var18, var19, var20, var21, var22), names_sep = "unique", names_repair = "universal")

duration_transit <- duration_transit %>%
                    select(contains("value"))

duration_transit <- duration_transit %>% 
                    pivot_longer(cols = starts_with("var"),
                                names_to = "route",
                                names_prefix = "wk",
                                values_to = "duration_seconds",
                                values_drop_na = FALSE
                              )

duration_transit$duration_hours <- duration_transit$duration_seconds / 3600

ggplot(duration_transit) + 
  geom_histogram(aes(duration_hours))

summary(duration_transit$duration_hours)
```

### Combining distance and duration for transit

```{r}
distance_duration_transit <- full_join(distance_transit, duration_transit)
```

# DRIVING: Getting driving routes and points

```{r}
car_directions <- apply(points_sample, 1, function(x){
  res <- google_directions(
    origin = x[['lat_lon']], 
    destination = x[['destination']],
    arrival_time =  arrival_time,
    mode = "driving",
    alternatives = FALSE,
    units = "metric"
  )
})
```

### Checking geocode and route status

```{r}
get_elem(car_directions, "status", recursive = TRUE, DF.as.list = TRUE)
```

### Getting the distance data from car

```{r}
distance_car <- get_elem(car_directions, "distance", recursive = TRUE, DF.as.list = TRUE)
distance_car <- as_tibble(distance_car)
distance_car <- distance_car %>% 
                            rename_with(~paste0("var", seq_along(.)))
distance_car <- distance_car %>%
                            filter(!row_number() %in% c(2))

distance_car <- distance_car %>% unnest(c(var1, var2, var3, var4, var5, var6, var7, var8, var9, var10, var11, var12, var13, var14, var15, var16, var17, var18, var19, var20, var21, var22), names_sep = "unique", names_repair = "universal")

distance_car <- distance_car %>%
                    select(contains("value"))

distance_car <- distance_car %>% 
                    pivot_longer(cols = starts_with("var"),
                                names_to = "route",
                                names_prefix = "wk",
                                values_to = "distance_meters",
                                values_drop_na = FALSE
                              )

distance_car$distance_km <- distance_car$distance_meters / 1000

ggplot(distance_car) + 
  geom_histogram(aes(distance_km))

summary(distance_car$distance_km)
```

### Getting duration data for car

```{r}
duration_car <- get_elem(car_directions, "duration", recursive = TRUE, DF.as.list = TRUE)
duration_car <- as_tibble(duration_car)
duration_car <- duration_car %>% 
                            rename_with(~paste0("var", seq_along(.)))
duration_car <- duration_car %>%
                            filter(!row_number() %in% c(2))

duration_car <- duration_car %>% unnest(c(var1, var2, var3, var4, var5, var6, var7, var8, var9, var10, var11, var12, var13, var14, var15, var16, var17, var18, var19, var20, var21, var22), names_sep = "unique", names_repair = "universal")

duration_car <- duration_car %>%
                    select(contains("value"))

duration_car <- duration_car %>% 
                    pivot_longer(cols = starts_with("var"),
                                names_to = "route",
                                names_prefix = "wk",
                                values_to = "duration_seconds",
                                values_drop_na = FALSE
                              )

duration_car$duration_hours <- duration_car$duration_seconds / 3600

ggplot(duration_car) + 
  geom_histogram(aes(duration_hours))

summary(duration_car$duration_hours)
```

### Combining distance and duration for car

```{r}
distance_duration_car <- full_join(distance_car, duration_car)
```

# BIKE+TRANSIT: Getting bike+transit routes and points

```{r}
points_sample_bt <- points_sample %>% select(lat_lon)
points_sample_bt$station <- origin     

bike_transit_directions <- apply(points_sample_bt, 1, function(x){
  res <- google_directions(
    origin = x[['lat_lon']], 
    destination = x[['station']],
    arrival_time =  arrival_time,
    mode = "bicycling",
    alternatives = FALSE,
    units = "metric"
  )
})
```

### Checking geocode and route status

```{r}
get_elem(bike_transit_directions, "status", recursive = TRUE, DF.as.list = TRUE)
```

### Getting the distance data for bike+transit (bike portion)

```{r}
distance_bike_transit <- get_elem(bike_transit_directions, "distance", recursive = TRUE, DF.as.list = TRUE)
distance_bike_transit <- as_tibble(distance_bike_transit)
distance_bike_transit <- distance_bike_transit %>% 
                            rename_with(~paste0("var", seq_along(.)))
distance_bike_transit <- distance_bike_transit %>%
                            filter(!row_number() %in% c(2))

distance_bike_transit <- distance_bike_transit %>% unnest(c(var1, var2, var3, var4, var5, var6, var7, var8, var9, var10, var11, var12, var13, var14, var15, var16, var17, var18, var19, var20, var21, var22), names_sep = "unique", names_repair = "universal")

distance_bike_transit <- distance_bike_transit %>%
                      select(contains("value"))

distance_bike_transit <- distance_bike_transit %>% 
                    pivot_longer(cols = starts_with("var"),
                                names_to = "route",
                                names_prefix = "wk",
                                values_to = "bike_to_station_distance_meters",
                                values_drop_na = FALSE
                              )

distance_bike_transit$bike_to_station_distance_km <- distance_bike_transit$bike_to_station_distance_meters / 1000

ggplot(distance_bike_transit) + 
  geom_histogram(aes(bike_to_station_distance_km))

summary(distance_bike_transit$bike_to_station_distance_km)
```

### Getting duration data for bike+transit (bike portion)

```{r}
duration_bike_transit <- get_elem(bike_transit_directions, "duration", recursive = TRUE, DF.as.list = TRUE)
duration_bike_transit <- as_tibble(duration_bike_transit)
duration_bike_transit <- duration_bike_transit %>% 
                            rename_with(~paste0("var", seq_along(.)))
duration_bike_transit <- duration_bike_transit %>%
                            filter(!row_number() %in% c(2))

duration_bike_transit <- duration_bike_transit %>% unnest(c(var1, var2, var3, var4, var5, var6, var7, var8, var9, var10, var11, var12, var13, var14, var15, var16, var17, var18, var19, var20, var21, var22), names_sep = "unique", names_repair = "universal")

duration_bike_transit <- duration_bike_transit %>%
                    select(contains("value"))

duration_bike_transit <- duration_bike_transit %>% 
                    pivot_longer(cols = starts_with("var"),
                                names_to = "route",
                                names_prefix = "wk",
                                values_to = "bike_to_station_duration_seconds",
                                values_drop_na = FALSE
                              )

duration_bike_transit$bike_to_station_duration_hours <- duration_bike_transit$bike_to_station_duration_seconds / 3600

ggplot(duration_bike_transit) + 
  geom_histogram(aes(bike_to_station_duration_hours))

summary(duration_bike_transit$bike_to_station_duration_hours)
```

## Bike+Transit (Transit portion)

```{r}
transit_bike <- google_directions(origin = origin,  
                          destination = destination,             
                          arrival_time =  arrival_time,
                          mode = "bicycling",
                          alternatives = FALSE,
                          units = "metric"
                          )
```

```{r}
distance_transit_bike <- get_elem(transit_bike, "distance", recursive = TRUE, DF.as.list = TRUE)
distance_transit_bike <- as_tibble(distance_transit_bike)
distance_transit_bike <- distance_transit_bike %>% slice(1)
distance_transit_bike <- distance_transit_bike$distance
distance_transit_bike <- distance_transit_bike %>% select(contains("value"))
distance_transit_bike <- as.vector(distance_transit_bike$value)
```

```{r}
duration_transit_bike <- get_elem(transit_bike, "duration", recursive = TRUE, DF.as.list = TRUE)
duration_transit_bike <- as_tibble(duration_transit_bike)
duration_transit_bike <- duration_transit_bike %>% slice(1)
duration_transit_bike <- duration_transit_bike$duration
duration_transit_bike <- duration_transit_bike %>% select(contains("value"))
duration_transit_bike <- as.vector(duration_transit_bike$value)
```


### Combining distance and duration for transit

```{r}
distance_duration_transit__bike <- full_join(distance_bike_transit, duration_bike_transit)

#### Distance
distance_duration_transit__bike$transit_to_location_distance_meters <- distance_transit_bike
distance_duration_transit__bike$transit_to_location_distance_km <- distance_duration_transit__bike$transit_to_location_distance_meters / 1000

distance_duration_transit__bike$distance_km <- distance_duration_transit__bike$bike_to_station_distance_km + distance_duration_transit__bike$transit_to_location_distance_km

distance_duration_transit__bike$distance_meters <- distance_duration_transit__bike$bike_to_station_distance_meters + distance_duration_transit__bike$transit_to_location_distance_meters

#### Duration
distance_duration_transit__bike$transit_to_location_duration_seconds <- duration_transit_bike
distance_duration_transit__bike$transit_to_location_duration_hours <- distance_duration_transit__bike$transit_to_location_duration_seconds / 3600

distance_duration_transit__bike$duration_hours <- distance_duration_transit__bike$bike_to_station_duration_hours + distance_duration_transit__bike$bike_to_station_duration_hours

distance_duration_transit__bike$duration_seconds <- distance_duration_transit__bike$bike_to_station_duration_seconds + distance_duration_transit__bike$bike_to_station_duration_seconds

distance_duration_transit_bike <- select(distance_duration_transit__bike, route, distance_meters, duration_hours, distance_km, duration_hours, duration_seconds)
```

# Combining bike, transit, and car data

```{r}
distance_duration_bike$mode <- "bike"
distance_duration_transit$mode <- "transit"
distance_duration_car$mode <- "car"
distance_duration_transit_bike$mode <- "bike_transit"


full_data1 <- full_join(distance_duration_bike, distance_duration_transit)
full_data2 <- full_join(full_data1, distance_duration_car)
full_data <- full_join(full_data2, distance_duration_transit_bike)

full_data$station <- origin
full_data$destination <- destination
```

# Writing CSV Files

```{r}
write_csv(distance_duration_transit__bike, file = "ottawa_barrhaven_bike_plus_transit.csv")    #### EDIT ME
write_csv(full_data, file = "ottawa_barrhaven_bike_transit_drive.csv")  #### EDIT ME
```

```{r, include = FALSE}
clear_keys()
```

